---
title: "247-commit-scraper"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

# Overview

This notebook collects commit data from the 247 team commit pages. The commit data will provide several metrics present within our commit likelihood model.

# Data Collection

Load relevant packages
```{r}
library(foreach)
library(doParallel)
library(tidyverse)
library(rvest)
library(assertr)
```

Set up parallelization
```{r warning = FALSE}
myCluster <- makeCluster(detectCores()-1, type = "PSOCK")
registerDoParallel(myCluster)
```

Function to get info from school's commit page for a specific year
```{r}
page_info <- function(page){
  #number of total commits for a school
  n_commits <- page %>% html_node(".ri-page__heading span") %>% html_text() %>% gsub("[^0-9]", "", .) %>% 
    as.numeric() %>% ifelse(is.na(.), 0, .)
  #number of total transfers for a school
  n_transfers <- page %>% html_element(".portal-list_itm .name") %>% html_text() %>% gsub("[^0-9]", "", .) %>% 
    as.numeric() %>% ifelse(is.na(.), 0, .)
  #commit player data
  commits <- page %>% html_elements(".wrapper")
  #transfer player data
  transfers <- page %>% html_elements(".portal-list_itm+ .portal-list_itm , .list-header+ .portal-list_itm")
  #save each element of the info page in a list
  output <- list(n_commits, commits, n_transfers, transfers)
  
  #return the output list
  return(output)
}
```

Function to get commit info into a clean dataframe
```{r}
commit_metrics <- function(commits){
  #player url
  link <- commits %>% html_element("a") %>% html_attr("href") %>% str_c("https:", .)
  #player name
  name <- commits %>% html_element("a") %>% html_text()
  #player school and city
  school_city <- commits %>% html_element("span") %>% html_text(trim = T)
  #player position
  position <- commits %>% html_element(css = "[class = position]") %>% html_text(trim = T)
  #player height and weight
  height_weight <- commits %>% html_element(css = "[class = metrics]") %>% html_text(trim = T)
  #number of stars
  stars <- commits %>% html_element(css = "[class = ri-page__star-and-score]") %>% 
    sapply(function(element){element %>% str_count("yellow")}) %>% suppressWarnings()
  #composite rating
  composite_rating <- commits %>% html_element(css = "[class = ri-page__star-and-score]") %>% html_text(trim = T)
  #national rank
  national_rank <- commits %>% html_nodes("a.natrank") %>% html_text()
  #position rank
  position_rank <- commits %>% html_nodes("a.posrank") %>% html_text()
  #state rank
  state_rank <- commits %>% html_nodes("a.sttrank") %>% html_text()
  #school that player ended up committing to
  status <- commits %>% html_nodes("div.status") %>% html_element("p.commit-date") %>% html_text(trim = T)
  #player level
  player_level <- rep("commit", length(commits))
  #combine all player info into dataframe
  commit_metrics_df <- cbind(link, name, school_city, position, height_weight, 
                             stars, composite_rating, 
                             national_rank, position_rank, state_rank, status, player_level) %>% as.data.frame()
  #return output dataframe
  return(commit_metrics_df)
}
```

Function to get transfer info into a clean dataframe
```{r}
transfer_metrics <- function(transfers){
  #player url
  link <- transfers %>% html_element("a") %>% html_attr("href")
  #player name
  name <- transfers %>% html_element("a") %>% html_text()
  #player position
  position <- transfers %>% html_element(css = "[class = position]") %>% html_text(trim = T)
  #player height and weight
  height_weight <- transfers %>% html_nodes("div.metrics") %>% html_text(trim = T)
  #number of stars
  stars <- transfers %>% html_element("div.rating") %>% html_element("span") %>% 
    sapply(function(element){element %>% str_count("yellow")}) %>% suppressWarnings()
  #composite rating
  composite_rating <- transfers %>% html_element("div.rating") %>% html_element("span") %>% 
    html_element("span.score") %>% html_text() %>% str_split(" ") %>% lapply(`[`, 1) %>% unlist()
  #rating level (HS/College)
  rating_level <- transfers %>% html_element("div.rating") %>% html_element("span") %>% 
    html_element("span.score") %>% html_text() %>% str_split(" ") %>% lapply(`[`, 2) %>% unlist() %>% 
    gsub("[^[:alnum:] ]", "", .)
  #eligibility
  eligibility <- transfers %>% html_element("div.eligibility") %>% html_text(trim = T)
  #former school
  former_school <- transfers %>% html_elements("div.transfer-institution") %>% html_element("img") %>% html_attr("alt")
  #former school link
  former_school_link <- transfers %>% html_elements("div.transfer-institution") %>% html_element("img") %>% 
    html_attr("src") %>% str_split("\\?") %>% lapply(`[`, 1) %>% unlist()
  #player level
  player_level <- rep("transfer", length(transfers))
  #combine all player info into dataframe
  transfer_metrics_df <- cbind(link, name, position, height_weight,
                               stars, composite_rating, rating_level,
                               eligibility, former_school, former_school_link, player_level) %>% as.data.frame()
  #return output dataframe
  return(transfer_metrics_df)
}
```

Initial framework for grabbing both ratings for transfer players (we stuck with most recent for simplicity's sake)
If used, would be included within the transfer_metrics function
```{r}
# transfer_rating <- transfers %>% html_element("div.rating") %>% html_elements("span")
# 
# transfers %>% html_element("div.rating") %>% html_element("span") %>% 
#   sapply(function(element){element %>% str_count("yellow")}) %>% suppressWarnings()
# 
# transfers %>% html_element("div.rating") %>% 
#   sapply(function(element){element %>% str_count("span")/16}) %>% suppressWarnings() %>% cumsum()
# 
# transfer_rating[seq(1, length(transfer_rating), 8)] %>% 
#   sapply(function(element){element %>% str_count("yellow")}) %>% suppressWarnings()
# transfer_rating[seq(1, length(transfer_rating), 8)] %>% html_element("span.score") %>% html_text() %>% 
#   str_split(" ") %>% lapply(`[`, 1) %>% unlist()
# transfer_rating[seq(1, length(transfer_rating), 8)] %>% html_element("span.score") %>% html_text() %>% 
#   str_split(" ") %>% lapply(`[`, 2) %>% unlist() %>% gsub("[^[:alnum:] ]", "", .)
```

Function to merge commit and transfer data 
```{r}
data_merge <- function(commit_data, transfer_data, relevant_cols) {
  #stack the commit and transfer data only storing the relevant columns
  output <- rbind(commit_data[,relevant_cols], transfer_data[,relevant_cols]) %>% as.data.frame()
  #return the output dataframe
  return(output)
}
```

Function to combine data for all shown and hidden players
```{r}
page_data <- function(page_info, year, relevant_cols = c("link", "position", "stars", "composite_rating", "player_level")){
  
  #exit function if no commits/transfers in the given year
  stopifnot(page_info[[1]] + page_info[[3]] > 0)
  
  #create dataframe for both commits and transfers, if/else allows for error handling of empty dataframes for either one
  if(page_info[[1]]==0) {
    all_commits <- transfer_metrics(page_info[[4]])[,relevant_cols]
    } else if(page_info[[3]]==0) {
    all_commits <- commit_metrics(page_info[[2]])[,relevant_cols]
    } else {
    all_commits <- data_merge(commit_data = commit_metrics(page_info[[2]]), 
                              transfer_data = transfer_metrics(page_info[[4]]), 
                              relevant_cols)
  }
  
  #verify the number of total commits matches scraped data, append column for year, save as output dataframe
  output <- all_commits %>% verify(nrow(.) == page_info[[1]] + page_info[[3]]) %>% mutate(year = year) %>% 
    select(year, everything())
  
  #return output
  return(output)
}
```

Function to get commit data for all years for a single team
```{r}
team_data <- function(teams, t, years) {
  #initialize the current team link
  team_link <- teams$link[t]
  #initialize the current team name
  team_name <- teams$name[t]
  #initialize empty df for all commits
  school_df <- data.frame()

  #loop over years
  for (i in years) {
  
    # initialize url, try to access the page
    url <- paste0(team_link, "Season/", i, "-Football/Commits/")
    #read the page html
    page <- read_html(url)
    #skip to next year if page does not exist
    if (is.na(page)) {next}
    #set variable to skip to next year for error handling
    skip_to_next <- F
    #tryCatch handler to get commit data for that year or skip to next iteration if an error is encountered 
    #(ie no commits that year, invalid url, etc.)
    output <- tryCatch(page_data(page_info(page), i), error = function(e) {skip_to_next <<- T})
    #skip to next iteration if there is an error
    if (skip_to_next==T) {next}
    #append results to previously initialized df for school
    school_df <- rbind(school_df, output)
  }
  #rework column names and order for output df
  school_df <- school_df %>% mutate(school = team_name) %>% select(school, everything())
  #return output df
  return(school_df)
}
```

Parallel implementation with time check
```{r}
#system time records the amount of time it takes to run
system.time({
  #initialize teams and years to be included
  teams <- read_csv("data/247-team.csv")
  years <- 1999:2021
  #run loop parallelized on several clusters and save into dataframe
  all_df <- foreach (t = 1:nrow(teams), .combine = "rbind", .packages = c("tidyverse", "rvest", "assertr")) %dopar%
    team_data(teams = teams, t = t, years = years)  
})
```

Stop cluster
```{r}
stopCluster(myCluster)
```


# Post-Data Collection Investigation and Cleaning

Display teams that we don't have commits for. Upon closer inspection, 247 does not list a unique team link for these schools.
```{r}
data.frame(teams$name) %>% anti_join(data.frame(all_df$school), by = c("teams.name"="all_df.school"))
```

Save output file (full data cleaning is outside of the scope of this document)
```{r}
output <- all_df %>% 
  mutate(composite_rating = ifelse(composite_rating=="NA", NA, composite_rating)) %>% 
  rename(committed = school, class = year) %>% 
  select(link, everything())
```

# Data Export

Export the output dataframe to a csv file for later use
```{r}
path <- paste0(getwd(), "/data/")
write_csv(output, file.path(path, "247-team-commit.csv"))
```
