---
title: "247-offer-scraper"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

# Overview

This notebook collects offer data from the 247 team offers pages. The point of this data is to support offer scraping in future notebooks. The offer data will provide several metrics present within our commit likelihood model.

# Data Collection

Load relevant packages
```{r}
library(foreach)
library(doParallel)
library(tidyverse)
library(rvest)
library(assertr)
```

Set up parallelization
```{r warning = FALSE}
myCluster <- makeCluster(detectCores()-1, type = "PSOCK")
registerDoParallel(myCluster)
```

Function to get info from school's offer page for a specific year
```{r}
page_info <- function(page){
  #number of total offers for a school
  n_offers <- page %>% html_node(".ri-page__heading span") %>% html_text() %>% gsub("[^0-9]", "", .) %>% as.numeric()
  #players that are displayed
  shown_players <- page %>% html_elements(".wrapper")
  #links to display players which are hidden bc 247 sucks
  hidden_position_links <- page %>% html_elements(".showmore_blk") %>% 
    html_element("a") %>% html_attr("href")
  
  #clean up hidden position links
  if(length(hidden_position_links) > 0){
    hidden_position_links = hidden_position_links %>% str_replace_all(" ", "%20") %>% str_c("https://247sports.com", .)
  }
  
  #save each element of the info page in a list
  output <- list(n_offers, shown_players, hidden_position_links)
  
  #return the output list
  return(output)
}
```


Function to get player info into a clean dataframe
```{r}
offer_metrics <- function(players){
  #player url
  link <- players %>% html_element("a") %>% html_attr("href") %>% str_c("https:", .)
  #player name
  name <- players %>% html_element("a") %>% html_text()
  #player school and city
  school_city <- players %>% html_element("span") %>% html_text(trim = T)
  #player position
  position <- players %>% html_element(css = "[class = position]") %>% html_text(trim = T)
  #player height and weight
  height_weight <- players %>% html_element(css = "[class = metrics]") %>% html_text(trim = T)
  #number of stars
  stars <- players %>% html_element(css = "[class = ri-page__star-and-score]") %>% 
    sapply(function(element){element %>% str_count("yellow")}) %>% suppressWarnings()
  #composite rating
  composite_rating <- players %>% html_element(css = "[class = ri-page__star-and-score]") %>% html_text(trim = T)
  #national rank
  national_rank <- players %>% html_nodes("a.natrank") %>% html_text()
  #position rank
  position_rank <- players %>% html_nodes("a.posrank") %>% html_text()
  #state rank
  state_rank <- players %>% html_nodes("a.sttrank") %>% html_text()
  #school that player ended up committing to
  final_commit <- players %>% html_nodes("div.status") %>% html_element("a") %>% html_element("img") %>% html_attr("alt")
  #image for school that player ended up committing to
  final_commit_image <- players %>% html_elements("div.status") %>% html_element("img") %>% html_attr("data-src") %>% 
    str_split("\\?") %>% lapply(`[`, 1) %>% unlist()
  #combine all player info into dataframe
  offer_metrics_df <- cbind(link, name, school_city, position, height_weight, 
                            stars, composite_rating, 
                            national_rank, position_rank, state_rank, final_commit, final_commit_image) %>% as.data.frame()
  #return output dataframe
  return(offer_metrics_df)
}
```

Function to combine data for all shown and hidden players
```{r}
page_data <- function(page_info, year){
  
  #exit function if no offers in the given year
  stopifnot(page_info[[1]] > 0)
  
  #player data for shown players into dataframe
  offer_df <- offer_metrics(page_info[[2]])
  
  #player data for hidden players appended to created dataframe
  for(i in page_info[[3]]) {
    hidden_players <- read_html(i) %>% html_elements(".wrapper")
    offer_df <- rbind(offer_df, offer_metrics(hidden_players))
  }
  
  #return dataframe and verify that the number of players equals the number of offers displayed on the page
  return(offer_df %>% verify(nrow(.) == page_info[[1]]) %>% mutate(year = year) %>% 
    select(year, everything()))
}
```

Function to get offer data for all years for a single team
```{r}
team_data <- function(teams, t, years) {
  #initialize the current team link
  team_link <- teams$link[t]
  #initialize the current team name
  team_name <- teams$name[t]
  #initialize empty df for all offers
  school_df <- data.frame()

  #loop over years
  for (i in years) {
  
    # initialize url, try to access the page
    url <- paste0(team_link, "Season/", i, "-Football/Offers/")
    #read the page html
    page <- read_html(url)
    #skip to next year if page does not exist
    if (is.na(page)) {next}
    #set variable to skip to next year for error handling
    skip_to_next <- F
    #tryCatch handler to get offer data for that year or skip to next iteration if an error is encountered 
    #(ie no offers that year, invalid url, etc.)
    output <- tryCatch(page_data(page_info(page), i), error = function(e) {skip_to_next <<- T})
    #skip to next iteration if there is an error
    if (skip_to_next==T) {next}
    #append results to previously initialized df for school
    school_df <- rbind(school_df, output)
  }
  #rework column names and order for output df
  school_df <- school_df %>% mutate(school = team_name) %>% select(school, everything())
  #return output df
  return(school_df)
}
```

Parallel implementation with time check
```{r}
#system time records the amount of time it takes to run
system.time({
  #initialize teams and years to be included
  teams <- read_csv("data/247-team.csv")
  years <- 1999:2021
  #run loop parallelized on several clusters and save into dataframe
  all_df <- foreach (t = 1:nrow(teams), .combine = "rbind", .packages = c("tidyverse", "rvest", "assertr")) %dopar%
    team_data(teams = teams, t = t, years = years)  
})
```

Stop cluster
```{r}
stopCluster(myCluster)
```

# Post-Data Collection Investigation and Cleaning

Display teams that we don't have offers for. Upon closer inspection, 247 does not list a unique team link for these schools.
```{r}
data.frame(teams$name) %>% anti_join(data.frame(all_df$school), by = c("teams.name"="all_df.school"))
```

Save output file (full data cleaning is outside of the scope of this document)
```{r}
output <- all_df %>% 
  mutate(final_commit_image = ifelse(final_commit_image=="", NA, final_commit_image)) %>% 
  rename(offer = school, class = year) %>% 
  select(link, everything())
```

# Data Export

Export the output dataframe to a csv file for later use
```{r}
path <- paste0(getwd(), "/data/")
write_csv(output, file.path(path, "247-team-offer.csv"))
```
